---
title: "First evaluation of alate aphids (16.05.2018)"
author: "Niemann, Menssen, Poehling (2020)"
output:
  pdf_document: default
  word_document: default
editor_options:
  chunk_output_type: console
urlcolor: blue
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache=TRUE, message=FALSE, error=FALSE, warning=FALSE)

```

```{r, echo=FALSE, message=FALSE}

# Loading some packages

# Data management and plots
library(tidyverse)

# Multiple comparisons and confidence intervals
library(emmeans)

# Compact letter display
library(multcompView)


# For mixed model
library(lme4)
library(lmerTest)

```


```{r, echo=FALSE}
# setwd("C:/Users/Menssen/Desktop/Beratung/niemann_jan_uwe/neg_bin_counts")

dat_raw <- read.csv2("Bonitur_1final.csv", header=TRUE, strip.white=TRUE)

dat_raw$Plot <- factor(dat_raw$P)
dat_raw$Reihe <- factor(dat_raw$Prow)
dat_raw$Reihe_Plot <- dat_raw$Reihe:dat_raw$Plot
dat_raw$inpl <- factor(dat_raw$inpl)
dat_raw$Pnr <- factor(dat_raw$Pnr)


# str(dat)
```



```{r,  echo=FALSE, message=FALSE}
# Building sums
datsum <- dat_raw %>% 
        dplyr::select(-(rep)) %>%
        group_by(Var, Plot, inpl) %>% 
        summarize(aphids_sum=sum(alaph),
                  n=n(),
                  scaled_sum=((aphids_sum+1)/n)) %>% 
  rename("var"=Var, 
         "plot"=Plot)
  
```


### Data
The statistical analysis was run on the following data set called `datsum` with the variables:

* _var:_ Different treatments
* _plot:_ Plots on the field
* _inpl:_ Location of the lettuce plants in the plots (0: border, 1: center )
* _aphids_sum:_  sum of flying aphids per location in each plot
* _n:_ Number of lettuce plants for each sum
* _scaled_sum:_ scaled sums calculated as (aphids_sum+1)/n 

```{r, echo=FALSE, tab.cap="First rows of datsum"}

# datsum <- data.frame(datsum)
datsum %>% 
  mutate(scaled_sum=round(scaled_sum, 3)) %>% 
  knitr::kable(caption="Summed total number of aphids")
```

\newpage

### Statistical analysis
A linear mixed model was fit to the data as described in the text

```{r,  eval=FALSE, message=FALSE}
library(lme4)

lmm_fit <- lmer(log(scaled_sum) ~ var * inpl + (1|plot), data=datsum)
```

```{r, echo=FALSE, message=FALSE}

lmm_fit <- lmer(log(scaled_sum) ~ var * inpl + (1|plot), data=datsum)

```


### ANOVA
```{r, eval=FALSE}
library(lmerTest)

anova(lmm_fit)
```

```{r, echo=FALSE}


ano_df <- anova(lmm_fit) %>% 
  data.frame() %>% 
  rename("SSQ"=Sum.Sq,
         "MSQ"=Mean.Sq,
         "F_value"=F.value,
         "p_Value"=Pr..F.) %>% 
  mutate(SSQ=round(SSQ, 3),
         MSQ=round(MSQ, 3),
         F_value=round(F_value, 3),
         p_Value=round(p_Value, 3))
  
Effect <- c("var", "inpl", "var:inpl")


ano_df <- cbind(Effect, ano_df)

ano_df %>% 
  knitr::kable(caption="ANOVA table")
```

The ANOVA (table 2) revealed the significance ($\alpha=0.05$) of the treatment (Var) 
since the p-value is 0.002, but no location effect (inpl) nor a 
treatment-location interaction (Var:inpl). 

\newpage

### Mean comparisons between the treatments
The mean comparisons were run on the logarithm of the scaled sums (as the model does),
but are already back transformed to the original scale for easier interpretation.
The least square means of the scaled sums for each treatment, their standard error
and their 95% confidence intervals are given in table 3. 


```{r, eval=FALSE, message=FALSE, warning=FALSE}

# Model based least square means and ther comparisons
library(emmeans)
lmm_comp <- emmeans(lmm_fit, specs="var", contr="pairwise", type="response")

```

```{r, eval=FALSE, message=FALSE, warning=FALSE}
# LS-means for the treatments
lmm_comp$emmeans
```

```{r, echo=FALSE, , message=FALSE, warning=FALSE}
lmm_comp <- emmeans(lmm_fit, specs="var", contr="pairwise", type="response")
as.data.frame(lmm_comp$emmeans) %>%
        rename("mean"=response,
               "se"=SE,
               "lower"=lower.CL,
               "upper"=upper.CL) %>% 
        mutate(mean=round(mean, 3),
               se=round(se, 3),
               lower=round(lower, 3),
               upper=round(upper, 3)) %>% 
        knitr::kable(caption="Least square means and their confidence intervals")
```


The mean comparisons (contrasts) were run as differences on the log-scale. Therefore
back transformation results in the ratio between the means of the scaled sums These
ratios, their standard error and the corresponding p-values are given in table 4.
```{r, eval=FALSE}
# Contrast tests
lmm_comp$contrasts
```


```{r, echo=FALSE, message=FALSE}
# lmm_comp <- emmeans(lmm_fit, specs="var", type="response", contr="pairwise")

# contr <- lmm_comp$contrasts

as.data.frame(lmm_comp$contrasts) %>% 
        rename("se"=SE,
               "t_ratio"=t.ratio,
               "p_value"=p.value) %>% 
        mutate(ratio=round(ratio, 3),
               se=round(se, 3),
               t_ratio=round(t_ratio, 3),
               p_value=round(p_value, 3)) %>% 
        knitr::kable(caption="Contrasts and p-values")
```

\newpage

### Location effect 
The model based average scaled sums for each location were compared on log-scale. 
Back-transformations results in a ratio between both sums (0: border plant, 1: inner plant).
Please note that the ratio is not significantly different from zero.
```{r, message=FALSE, warning=FALSE}

# Model based least square means and ther comparisons for the location effect 
lmm_comp_inpl <- emmeans(lmm_fit, specs="inpl",  
                         contr="pairwise", type="response")

```

```{r, eval=FALSE}
# Ratio between the mean scaled sums for each location 
lmm_comp_inpl$contrasts
```


```{r, echo=FALSE}
as.data.frame(lmm_comp_inpl$contrasts) %>% 
        rename("se"=SE,
               "t_ratio"=t.ratio,
               "p_value"=p.value) %>% 
        mutate(ratio=round(ratio, 3),
               se=round(se, 3),
               t_ratio=round(t_ratio, 3),
               p_value=round(p_value, 3)) %>% 
        knitr::kable(caption="Ratios of scaled sums between the locations")
```


### Location effect in each treatment

Ratios between the scaled sums of the border plants (0) and the inner plants (1) 
split by the treatments are given in Tab. 6 . Please note, that none of these 
ratios is significantly different from zero.

```{r, message=FALSE, warning=FALSE}

# Model based least square means and ther comparisons for the location effect 
# in each treatment
lmm_comp_inter <- emmeans(lmm_fit, specs="inpl", by="var", 
                         contr="pairwise", type="response")

```

```{r, eval=FALSE}
# Ratios
lmm_comp_inter$contrasts
```


```{r, echo=FALSE}
as.data.frame(lmm_comp_inter$contrasts) %>% 
        rename("se"=SE,
               "t_ratio"=t.ratio,
               "p_value"=p.value) %>% 
        mutate(ratio=round(ratio, 3),
               se=round(se, 3),
               t_ratio=round(t_ratio, 3),
               p_value=round(p_value, 3)) %>% 
        knitr::kable(caption="Ratios of scaled sums between the locations and p-values")
```



